---
title: "Network Exploration"
output: html_notebook
---

Interactive view of an organization's data sources and the data flowing between them. This pulls from a Google Sheet. The Google Sheets document used below is at https://docs.google.com/spreadsheets/d/14QZbaKMNqGAGQpOPRZDZeHdb9H8kb19Vbv1VBItvpPs/edit?usp=sharing. You will want to make a copy of that and modify it for your own needs, publish and make public (it can be unlisted), and grab the ID from the URL and updated the `googlesheets_id` value below.

## Basic Setup and Get the Data

```{r setup, message=FALSE}

# See http://kateto.net/network-visualization for some details. Ultimately, it's all hinging on
# the visNetwork package, which has a number of different functions for adjusting the actual visualization.

# Load libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse,         # Includes dplyr, ggplot2, and others; very key!
               visNetwork,        # For actually creating the visualization
               RColorBrewer,      # For generating a color palette
               googlesheets)      # For retrieving the underlying data from Google Sheets

# Get the ID for the sheet. You can pull this from the URL. The URL for the Google Shee used
# here is at https://docs.google.com/spreadsheets/d/14QZbaKMNqGAGQpOPRZDZeHdb9H8kb19Vbv1VBItvpPs/edit?usp=sharing.
# Make a copy of that and then replace the ID below. Note that the sheet need to be
# *both* Public (unlisted is fine) AND "Published to web"
googlesheets_id <- "14QZbaKMNqGAGQpOPRZDZeHdb9H8kb19Vbv1VBItvpPs"

# Now, actually get the key for that ID
googlesheets_key <- gs_key(googlesheets_id, lookup=FALSE)

# Get the data
links_raw <- gs_read(googlesheets_key, ws="Data Flows", verbose=FALSE)
nodes_raw <- gs_read(googlesheets_key, ws="Systems", verbose=FALSE)

```

## Data Munging / Transformation

```{r data-munging, message=FALSE}

# Convert to network graphing ready objects. This is one set of edges (the data feeds)
# and one set of nodes (the data sources)

# The "links" are the "edges" (the lines connecting the different data sources).
# This part of the script, essentially, renames the columns in the data table to
# match the visNetwork() function's 'edges' argument requirements later in the script.
# Of course, you could just name things that way in the underlying Google Sheet,
# but underlying Google Sheets are rarely designed with that much friendliness.
links <- links_raw %>% 

  # Select the columns that are actually going to be used
  select(from = From,
         to = To)

# The "nodes" are the actual data sources (the circles)
nodes <- nodes_raw %>% 
  select(id = System,
         data_type = `Data Type`,
         owner = Owner)

####################
# We'll size the nodes based on the number of data flows to/from them. The sizing
# could be anything (it could be something to do with the size of the actual system, 
# for instance), so this code can be tweaked accordingly.

# Get the total feeds going *to* each data source
nodes_to_links <- nodes %>%
  inner_join(links, by = c(id = "to")) %>% 
  group_by(id) %>% 
  summarise(total_links_to = n())

# Get the total feeds going *from* each data source
nodes_from_links <- nodes %>%
  inner_join(links, by = c(id = "from")) %>% 
  group_by(id) %>% 
  summarise(total_links_from = n())

# Combine those together -- adding *to* and *from*
nodes_link_counts <- nodes_to_links %>% 
  full_join(nodes_from_links) %>% 
  mutate(total_links_to = ifelse(is.na(total_links_to), 
                                 0, total_links_to)) %>%       # Replace NAs with 0s for future math
  mutate(total_links_from = ifelse(is.na(total_links_from), 
                                   0, total_links_from)) %>%   # Replace NAs with 0s for future math
  mutate(size = total_links_to + total_links_from) %>% 
  select(id, size)

# Join those back to the main nodes data
nodes <- nodes %>% 
  left_join(nodes_link_counts) %>% 
  filter(!is.na(size)) %>%             # Remove orphaned nodes
  mutate(size = 5 + size * 3)          # Scale the circle sizes (requires some fiddling)

# Remove unused data frames
rm(nodes_to_links, nodes_from_links, nodes_link_counts)

```

## Visualization

```{r visualization, message=FALSE}

# Add some more options
nodes$shape <- "dot"             # Just use circles
nodes$shadow <- FALSE            # Nodes will drop shadow
nodes$title <- nodes$id          # Node title
nodes$label <- nodes$id          # Node label
nodes$borderWidth <- 0.5         # Node border width
nodes$group <- nodes$data_type   # Use data type for the legend
nodes$color.border <- "gray80"

links$color <- "#BBBBBB"    # line color  
links$arrows <- "middle"    # arrows: 'from', 'to', or 'middle'
links$smooth <- TRUE        # should the edges be curved?
links$shadow <- FALSE       # edge shadow

# Generate the plot. The visGroups are a hack here. There's some risk that they won't
# wind up being quite right if a new category gets introduced.
viz_output <- visNetwork(nodes, links, width = "1024px", height = "768px") %>%
  visOptions(highlightNearest = TRUE,
             selectedBy = list(variable = "owner", main="Select System Owner")) %>% 
  
  visLayout(randomSeed = 5003, improvedLayout = TRUE) %>% 
  visLegend(position = "right")
  
  # visLegend(position = "right", zoom=FALSE, width=0.12, useGroups = FALSE)

viz_output

visSave(viz_output, "data_ecosystem_visualization.html", selfcontained = TRUE, background = "white")

```
